---
title: "Global Terrorism"
author: "Weijia Zhang, Yuanzhu Li, Yijing Lin"
date: "12/3/2018"
output: github_document
---
Global Terrorism Dataset  


leaflet package or shiny plot 
#Context
The Global Terrorism Database (GTD) is an open-source database including information on terrorism around the world from 1970 through 2017 (with annual updates planned for the future). Unlike many other event databases, the GTD includes systematic data on domestic as well as international terrorist incidents that have occurred during this time period and now includes more than 180,000 cases.

#Goals:
We are going to use the global terrorism dataset to visualize the occurrence of terrorism across the globe from 1970 to 2017. In particular, we aim to show:
1)frequency of terrorism in 2017 across the globe  - world map 
2)Pick 5-8 countries of interest, explore the changes across the time span 
3)what the distribution of different attack types (such as suicide, bombing) is. 
4)What the distribution of number of wounded is for each attack typeï¼What the distribution of the use of different attack types based on different terrorist groups is 
5)explore the correlation between attack type and number of killed



#Content
Geography: Worldwide
Time period: 1970-2017, except 1993
Unit of analysis: Attack
Variables: >100 variables on location, tactics, perpetrators, targets, and outcomes
Sources: Unclassified media articles (Note: Please interpret changes over time with caution. Global patterns are driven by diverse trends in particular regions, and data collection is influenced by fluctuations in access to media coverage over both time and place.)
 
#Data(variables):
Iyear: This field contains the year in which the incident occurred.
Country_txt: This field identifies the country or location where the incident occurred
Region_txt: This field identifies the region in which the incident occurred.
City: Name of the city, village, or town in which the incident occurred
Success: Success of a terrorist strike
Suicide: 1 = "Yes" The incident was a suicide attack. 0 = "No" There is no indication that the incident was a suicide attack.
Targsubtype1_txt: The more specific target category
Nperps: The total number of terrorists participating in the incident
Nwound: Number of confirmed non-fatal injuries to both perpetrators and victims.
Attacktype: types of attacks
Gname: name of the terrorist group (potentially be coded based on ideology)
 
Website:https://www.kaggle.com/START-UMD/gtd

Data: Global Terrorism Dataset  
```{r package intallation}


```

```{r data cleaning}
globalterro<-read.csv('data/415Project.csv')
library(tidyverse)
problems(globalterro)
globalterro <- globalterro%>%select(iyear,imonth,country_txt, region_txt, city, nkill, latitude, longitude, attacktype1_txt, success, nperps,  gname )%>%filter(!is.na(iyear), !is.na(country_txt),!is.na(nkill))#clean data before each section below too
colnames(globalterro)[colnames(globalterro)=="country_txt"]<-'country'
colnames(globalterro)[colnames(globalterro)=="iyear"]<-'year'
colnames(globalterro)[colnames(globalterro)=="attacktype1_txt"]<-'attack_type'
```









```{r q2}
#HPick 5-8 countries of interest, explore the changes across the time span 

#number of incidents
Num_I_world<-globalterro%>%group_by(year)%>%mutate(world = n())%>%ggplot(aes(year,world))+geom_line()+geom_point(shape=23)+labs(x='year',y='terrorist attacks')+theme_bw()
Num_I_world
Num_incident5<-globalterro%>%filter(year>= 1991, country=='Syria'|country=='Iraq'|country== 'Afghanistan'|country== 'United States'|country=='United Kingdom'|country=='Yemen')
#Num_incident5

Num_incident5%>%group_by(country,year)%>%summarise(terrorist_attacks=n())%>%ggplot(aes(year,terrorist_attacks, color = country))+geom_line()+geom_point()+theme_bw()#the plot shows the number of terrorism incidents occured in the five countries from 1970 to 2017

Num_incident_isr_pal<-globalterro%>%filter(country=='Israel'|country=='West Bank and Gaza Strip')
Num_incident_isr_pal%>%group_by(country,year)%>%summarise(terrorist_attacks=n())%>%ggplot(aes(year,terrorist_attacks, color = country))+geom_line()+geom_point()+theme_bw()


gt_incidentworldrank<-globalterro%>%group_by(country)%>%summarize(number_of_incidents=n(),deaths=sum(nkill))%>%arrange(desc(number_of_incidents))
#gt_incidentworldrank

gt_incidentworldrank_before2003<-globalterro%>%filter(year<2003)%>%group_by(country)%>%summarize(number_of_incidents=n(),deaths=sum(nkill))%>%arrange(desc(number_of_incidents))
#gt_incidentworldrank_before2003
gt_incidentworldrank_after2003<-globalterro%>%filter(year>=2003)%>%group_by(country)%>%summarize(number_of_incidents=n(),deaths=sum(nkill))%>%arrange(desc(number_of_incidents))



gt_incidents_top4<-globalterro%>%filter(country=='Iraq'|country=='Pakistan'|country=='Afghanistan'|country=='India')%>%group_by(country,year)%>%summarise(number_of_incidents=n())
#gt_incidents_top4
gt_incidents_top4%>%ggplot(aes(year,number_of_incidents, color = country))+geom_line()+geom_point()+theme_bw()

gt_incidents_top3_after2003<-globalterro%>%filter(year>=2003,country=='Iraq'|country=='Pakistan'|country=='Afghanistan')%>%group_by(country)%>%summarize(terrorist_attack_after2003=n())
#gt_incidents_top3_after2003

gt_incidents_top3_before2003<-globalterro%>%filter(year<2003,country=='Iraq'|country=='Pakistan'|country=='Afghanistan')%>%group_by(country)%>%summarize(terrorist_attack_before2003=n())
#gt_incidents_top3_before2003

gt_incidents_top3_comparison<-full_join(gt_incidents_top3_after2003, gt_incidents_top3_before2003, by = "country")
gt_incidents_top3_comparison

#number of fatalities 
gt_countries<-globalterro%>%filter(year>= 1991, country=='Syria'|country=='Iraq'|country== 'Afghanistan'|country== 'United States'|country=='United Kingdom'|country=='Yemen')%>%group_by(country, year)%>%summarize(fatalities=sum(nkill))
#gt_countries
gt_countries%>%ggplot(aes(year, fatalities, color=country))+geom_point()+geom_line()+theme_bw()#the plot shows the fatalities rates resulted from terrorism in six countries from 1970 to 2017

gt_911<-globalterro%>%filter(country=='Iraq'|country== 'Afghanistan'|country== 'United States')%>%group_by(country, year)%>%summarize(fatalities=sum(nkill))
gt_911%>%ggplot(aes(year, fatalities, color=country))+geom_point()+geom_line()+theme_bw()

gt_isr_pal<-globalterro%>%filter(country=='Israel'|country=='West Bank and Gaza Strip')%>%group_by(country, year)%>%summarize(fatalities=sum(nkill))
gt_isr_pal%>%ggplot(aes(year, fatalities, color=country))+geom_point()+geom_line()+labs(x='year')+theme_bw()
gt_fatalities_worldrank<-globalterro%>%group_by(country)%>%summarize(fatalities=sum(nkill))%>%arrange(desc(fatalities))
head(gt_fatalities_worldrank)#the table that contains the rank of countries based on the fatalities rates resulted from terrorism since 1970 to 2017



```


```{r q3 attack types}
number_of_fatalities<-sum(globalterro$nkill)
gt_attacktypes<-globalterro%>%filter(attack_type!='unknown')%>%group_by(attack_type)%>%summarize(fatalities=sum(nkill), rates=fatalities/number_of_fatalities,incidents=n())%>%arrange(desc(fatalities))
gt_attacktypes
gt_attacktypes%>%ggplot()+geom_bar(aes(x=rates,fill=attack_type))+coord_polar()+theme_bw()

#6 countries
number_of_fatalities_per_country<-globalterro%>%filter(country=='Iraq'|country== 'United States')%>%group_by(country,attack_type)%>%group_by(country)%>%summarize(fatalities=median(nkill))
number_of_fatalities_per_country
med_attack<-globalterro%>%group_by(attack_type)%>%summarize(avg=mean(nkill))
med_attack
gt_attacktypes_6countries<-globalterro%>%filter(country=='Iraq'|country== 'Afghanistan'|country== 'United States'|country=='France'|country=='Syria'|country=='Israel')%>%group_by(country,attack_type) #rates?
gt_attacktypes_6countries%>%ggplot()+geom_bar(aes(attack_type, fill=country),position='dodge')+theme_bw()

```

```{r q4 before and after 911 analysis }
globalterro911<-globalterro%>%mutate(idate=year*100+imonth)
#globalterro911
toptenc<-globalterro911%>%mutate(count_total=n())%>%group_by(country)%>%summarise(count=n())%>%arrange(desc(count))%>%head(10)
before911<- globalterro911%>%filter(idate <200109)%>%group_by(country)%>%summarise(cot_bf911=n())%>%arrange(desc(cot_bf911))
before911<-left_join(before911,toptenc,by="country")%>%mutate(num = count-cot_bf911)%>%mutate(before = cot_bf911/count)
#before911

after911<-globalterro911%>%filter(idate>=200109)%>%group_by(country)%>%summarise(cot_ps911=n())%>%arrange(desc(cot_ps911))
after911<-left_join(after911,toptenc,by = "country")%>%mutate(num = count-cot_ps911)%>%mutate(post = cot_ps911/count)
#after911
```




```{r 911}

toptenc<-globalterro911%>%mutate(count_total=n())%>%group_by(country)%>%summarise(count=n())%>%arrange(desc(count))%>%head(10)

globalterro911<-globalterro%>%mutate(idate=year*100+imonth)
#globalterro911
before911<- globalterro911%>%filter(idate <200109)%>%group_by(country)%>%summarise(cot_bf911=n())%>%arrange(desc(cot_bf911))
before911<-left_join(before911,toptenc,by="country")%>%mutate(num = count-cot_bf911)%>%mutate(before = cot_bf911/count)
before911topten <- before911%>% head(10)
before911topten 
after911<-globalterro911%>%filter(idate>=200109)%>%group_by(country)%>%summarise(cot_ps911=n())%>%arrange(desc(cot_ps911))
after911<-left_join(after911,toptenc,by = "country")%>%mutate(num = count-cot_ps911)%>%mutate(post = cot_ps911/count)
after911topeten <- after911%>% head(10)
after911topeten
compare <- before911topten %>%
   mutate(country_after = after911topeten$country)%>%mutate(count_post911 = after911topeten$cot_ps911)%>%mutate(total_count = after911topeten$count)%>%mutate(num_post = after911topeten$num)%>%mutate(post = after911topeten$post)
colnames(compare)[colnames(compare)=="country_after"]<-'country'
colnames(compare)[colnames(compare)=="cot_bf911"]<-'count_before'
colnames(compare)[colnames(compare)=="count_post911"] <- 'count_after'
colnames(compare)[colnames(compare)=="count"] <- 'total_count'
colnames(compare)[colnames(compare)=="before"] <- 'before_percentage'
colnames(compare)[colnames(compare)=="post"] <- 'after_percentage'
colnames(compare)[colnames(compare)=="num"] <- 'num_before'
library(knitr)
kable(compare)
```


```{r tmap}
library(tmap)
data("World")

combined1<-left_join(x=World,y=toptenc,by = c("name"="country"))
tm_shape(combined1)+tm_polygons("count", breaks = c(0,50,100,250,500,1000,2500))

combine_bf911 <- left_join(x=World, y=before911, by = c("name" = "country"))
tm_shape(combine_bf911)+tm_polygons("cot_bf911", breaks = c(0,50,100,250,500,1000,2500))

combine_af911 <- left_join(x=World, y=after911, by = c("name" = "country"))
tm_shape(combine_af911)+tm_polygons("cot_ps911", breaks = c(0,50,100,250,500,1000,2500))





```



